SAT Solving for Vesicle Traffic Systems in Cells

### 2017-07-20 paper/intro_MTedit.tex
done

The analysis of VTSs is a difficult computational problem because of the combinatorial scaling of graph topologies and regulatory rules. Many questions require us to check useful properties over all possible graphs and regulatory rules. Formal verification tools such as model checkers \cite{clarke1996symbolic, biere2003bounded, clarke2008birth, cimatti2000nusmv, holzmann1997model} and SAT solvers \cite{moskewicz2001chaff,een2004extensible} allow us to do this symbolically for graphs of finite size, without enumerating all instances. In a recent work, we employed the model checker CBMC \cite{CKY03, ckl2004} meant to analyze ANSI C programs for studying this problem. VTSs and SNARE regulations were modeled as C code using arrays to represent graphs and Boolean tables. Queries about graph connectivity requirements for different variations of SNARE regulations were modeled as logical assertions to be checked by the model checker. CBMC uses its own built-in encodings to reduce the analysis problem to a Boolean satisfiability problem which is solved using SAT solvers. However, the scalability of our recent approach was limited (up to VTSs with 8 nodes) due to the following reasons:

\begin{enumerate}
\item The encoding used to model VTSs was not fine-tuned to the structure of the problem.
\item The queries to be solved involved quantified formulas, i.e., we needed to solve queries of the form ``\emph{for all} graphs does there \emph{exist} SNARE regulation constraints that satisfy cellular steady state conditions.
\end{enumerate}

As CBMC does not directly support solving such quantified formulas, we had to use a technique that combined non-determinism and partial enumeration that is not as efficient.\newline

In this paper, we have developed a novel SAT encoding for searching a VTS that satisfy the given properties. We improved the encoding of some of the conditions for VTSs. In the earlier implementation, reachability was encoded by enumeration of paths. In our current encoding, we have recursively defined the reachability condition such that that we avoid the exponential enumeration. We have also encoded $k$-connectivity in a similar fashion and avoided a blowup in the formula size. Furthermore, our encoding supports several variations of the conditions. We have implemented the encoding using Z3~\cite{z3} python API and searched for VTS satisfying various properties upto size 14-18 nodes as compare earlier tool that could only search graph upto size 8 nodes. Our tool is useful to cell biologists since VTSs of real cells have approximately ten compartments.  \\

>>>>

The analysis of VTSs is a difficult computational problem because of the combinatorial scaling of graph topologies and regulatory rules. Many questions require us to check useful properties over all possible graphs and regulatory rules. Formal verification tools such as model checkers \cite{clarke1996symbolic, biere2003bounded, clarke2008birth, cimatti2000nusmv, holzmann1997model} and SAT solvers \cite{moskewicz2001chaff,een2004extensible} allow us to do this symbolically for graphs of finite size, without enumerating all instances. In a recent work, we employed the model checker CBMC \cite{CKY03, ckl2004} meant to analyze ANSI C programs for studying this problem. VTSs and SNARE regulations were modeled as C code using arrays to represent graphs and Boolean tables. Queries about graph connectivity requirements for different variations of SNARE regulations were modeled as logical assertions to be checked by the model checker. CBMC uses its own built-in encodings to reduce the analysis problem to a Boolean satisfiability problem which is solved using SAT solvers.
%
However, the scalability of our recent approach was limited (up to
VTSs with 6 nodes) due to the encoding used to model VTSs was not
fine-tuned to the structure of the problem.
%
In particular, we needed to encode conditional reachability between
nodes (refer section 3, stability condition), and solved using a
combination of non-determinism and partial enumeration that is not as
efficient.

% \begin{enumerate}
% \item 
% \item The \textbf{specification involved enumeration of all the possible paths in order to satisfy both the graph connectivity and the stability condition (refer section 3).}
% %quantified formulas, i.e., we needed to solve queries of the form ``\emph{for all} graphs does there \emph{exist} SNARE regulation constraints that satisfy cellular steady state conditions.
% \end{enumerate}

% \newline

% \todo{REQUIRE A REWRITE FLOW PROBS}
In this paper, we have developed a novel SAT encoding for
searching VTSs that satisfy the given properties.
%
We have improved the encoding of some of the conditions for VTSs.
%
Especially, we have recursively defined the reachability condition
such that that we avoid the inefficient enumeration.
%
% We have also encoded $k$-connectivity in a similar fashion and avoided
% a blowup in the formula size.
%
Furthermore, our encoding supports several variations of the properties.
%
We have implemented the encoding using Z3~\cite{z3} python API and
searched for VTS satisfying various properties upto size 14-18 nodes
as compare earlier tool that could only search graph upto size 6
nodes.
%
Our tool is useful to cell biologists since VTSs of real cells have
approximately ten compartments.

  % \\
